---
title: "Fast and parallelized math and stat functions"
author: "Philippe Grosjean (phgrosjean@sciviews.org)"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Fast and parallelized math and stat functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The {svFast} package provides a series of math and stat function that compute in parallel if the vector is large enough (by default >= 50,000 elements). Otherwise, these functions are simplified version of corresponding base R functions. They run on numeric vectors and do not retain any attributes of the input vector. They are **not** generic functions, on the contrary to their base R counterparts. They are designed to be fast when computing 100,000s to 100,000,000s items. Their names are the same as corresponding base functions followed by an underscore. For example, `log_()` is the fast version of `log()`.

```{r}
library(svFast)
# The number of threads used for calculation can be changed with:
#RcppParallel::setThreadOptions(numThreads = 4)
```

While vector `x` is short, the `fun_()` versions run sequentially at similar speed of base R equivalent for vectors roughly >= 1000 items. For smaller vectors, the overhead of Rcpp make these functions slower.

```{r}
x <- 1:10
log_(x)
identical(log_(x), log(x))
# log in base 8
log_(x, base = 8)
bench::mark(log(x), log_(x)) # Slower with such a short vector
```

When `x` is long, the `fun_()` versions run in parallel and are much faster than the base R equivalent.

```{r}
x2 <- runif(1e5)
# x2 size larger than 50,000, parallel computation activated
bench::mark(log(x2), log_(x2)) # Faster (depends on number of threads)
```
